---
title: "Scratchwork"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

```{r view-deaths-cases}
US_deaths_cases <- 
  read_csv("/cloud/project/data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
glimpse(US_deaths_cases)
```

```{r view-vaccinations}
US_vaccinations <- 
  read_csv("/cloud/project/data/COVID-19_Vaccinations_in_the_United_States_Jurisdiction.csv")
glimpse(US_vaccinations)
```

First we filter the data so it only includes the most up-to-date vaccination numbers as of 10/05/2021:

```{r october-vaccinations}
October_vaccinations <- filter(US_vaccinations, Date == "10/05/2021")
```

Next, we take the top 20 locations and create a bar plot to see which vaccines are in each respective location:

```{r vaccine-bar-chart, fig.width=10, fig.height=6}
October_vaccinations %>% 
  select(Location, Series_Complete_Yes, Series_Complete_Janssen,
         Series_Complete_Moderna, Series_Complete_Pfizer, 
         Series_Complete_Unk_Manuf) %>%
  filter(Location != "US") %>%
  group_by(Location) %>%
  arrange(desc(Series_Complete_Yes)) %>%
  head(20) %>%
  pivot_longer(cols = -Location, names_to = "Series_Type") %>%
  filter(Series_Type == "Series_Complete_Janssen" | 
           Series_Type == "Series_Complete_Moderna" |
           Series_Type == "Series_Complete_Pfizer" | 
           Series_Type == "Series_Complete_Unk_Manuf") %>%
  ggplot(aes(x = reorder(Location, value), 
             y = value/ 10^6, 
             fill = Series_Type)) +
  geom_bar(stat = "identity") + 
  labs(title = "US Vaccinations Breakdown",
       x = "Location",
       y = "People (in millions)",
       fill = "Type of Vaccine") +
  coord_flip() + 
  theme_bw() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20))
```
 
It looks like the Pfizer vaccine is the most popular across the board. If we want a better visualization to compare states, we could filter out the `US` location from this chart.

```{r vaccine-bar-chart2, fig.width=12, fig.height=10}
October_vaccinations %>% 
  select(Location, Series_Complete_Yes, Series_Complete_Janssen,
         Series_Complete_Moderna, Series_Complete_Pfizer, 
         Series_Complete_Unk_Manuf) %>%
  filter(!(Location %in% c("VI", "VA2", "DD2", 
                       "PR", "IH2", "BP2", 
                       "GU", "FM", "MP", 
                       "AS", "MH", "RP", 
                       "LTC"))) %>%
  group_by(Location) %>%
  arrange(desc(Series_Complete_Yes)) %>%
  pivot_longer(cols = -Location, names_to = "Series_Type") %>%
  filter(Series_Type == "Series_Complete_Janssen" | 
           Series_Type == "Series_Complete_Moderna" |
           Series_Type == "Series_Complete_Pfizer" | 
           Series_Type == "Series_Complete_Unk_Manuf") %>%
  ggplot(aes(x = reorder(Location, value), 
             y = value/ 10^6, 
             fill = Series_Type)) +
  geom_bar(stat = "identity", position = "fill") + 
  labs(title = "US Vaccinations Breakdown",
       x = "Location",
       y = "Fraction of Population",
       fill = "Type of Vaccine") +
  coord_flip() + 
  theme_bw() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) 
```

Vaccination breakdown for all 50 states and DC.

```{r}
new_US_deaths_cases <- US_deaths_cases %>%
  filter(!(state %in% c("NYC", "PR", "GU",
                           "VI", "MP", "RMI",
                           "AS", "PW", "FSM"))) %>%
  group_by(submission_date) %>%
  summarise(total_new_cases = sum(new_case),
            total_new_deaths = sum(new_death)) 

new_US_deaths_cases$submission_date = as.Date(new_US_deaths_cases$submission_date,
                                              format="%m/%d/%Y")

new_US_deaths_cases
```

```{r, fig.width=20, fig.height=12}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_new_cases)) +
  labs(title = "New Cases vs. Time",
       x = "Date",
       y = "Number of People") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r, fig.width=20, fig.height=12}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_new_deaths)) +
  labs(title = "New Deaths vs. Time",
       x = "Date",
       y = "Number of People") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

