---
title: "Scratchwork"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(lubridate)
library(devtools)
library(dsbox)
library(ggridges)
```

```{r view-deaths-cases}
US_deaths_cases <- 
  read_csv("/cloud/project/data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
glimpse(US_deaths_cases)
```

```{r view-vaccinations}
US_vaccinations <- 
  read_csv("/cloud/project/data/COVID-19_Vaccinations_in_the_United_States_Jurisdiction.csv")
glimpse(US_vaccinations)
```

First we filter the data so it only includes the most up-to-date vaccination numbers as of 10/05/2021:

```{r october-vaccinations}
October_vaccinations <- filter(US_vaccinations, Date == "10/05/2021")
```

Next, we take the top 20 locations and create a bar plot to see which vaccines are in each respective location:

```{r vaccine-bar-chart, fig.width=10, fig.height=6}
October_vaccinations %>% 
  select(Location, Series_Complete_Yes, Series_Complete_Janssen,
         Series_Complete_Moderna, Series_Complete_Pfizer, 
         Series_Complete_Unk_Manuf) %>%
  filter(Location != "US") %>%
  group_by(Location) %>%
  arrange(desc(Series_Complete_Yes)) %>%
  head(20) %>%
  pivot_longer(cols = -Location, names_to = "Series_Type") %>%
  filter(Series_Type == "Series_Complete_Janssen" | 
           Series_Type == "Series_Complete_Moderna" |
           Series_Type == "Series_Complete_Pfizer" | 
           Series_Type == "Series_Complete_Unk_Manuf") %>%
  ggplot(aes(x = reorder(Location, value), 
             y = value/ 10^6, 
             fill = Series_Type)) +
  geom_bar(stat = "identity") + 
  labs(title = "US Vaccinations Breakdown",
       x = "Location",
       y = "People (in millions)",
       fill = "Type of Vaccine") +
  coord_flip() + 
  theme_bw() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20))
```
 
It looks like the Pfizer vaccine is the most popular across the board. If we want a better visualization to compare states, we could filter out the `US` location from this chart.

```{r vaccine-bar-chart2, fig.width=12, fig.height=10}
October_vaccinations %>% 
  select(Location, Series_Complete_Yes, Series_Complete_Janssen,
         Series_Complete_Moderna, Series_Complete_Pfizer, 
         Series_Complete_Unk_Manuf) %>%
  filter(!(Location %in% c("VI", "VA2", "DD2", 
                       "PR", "IH2", "BP2", 
                       "GU", "FM", "MP", 
                       "AS", "MH", "RP", 
                       "LTC"))) %>%
  group_by(Location) %>%
  arrange(desc(Series_Complete_Yes)) %>%
  pivot_longer(cols = -Location, names_to = "Series_Type") %>%
  filter(Series_Type == "Series_Complete_Janssen" | 
           Series_Type == "Series_Complete_Moderna" |
           Series_Type == "Series_Complete_Pfizer" | 
           Series_Type == "Series_Complete_Unk_Manuf") %>%
  ggplot(aes(x = reorder(Location, value), 
             y = value/ 10^6, 
             fill = Series_Type)) +
  geom_bar(stat = "identity", position = "fill") + 
  labs(title = "US Vaccinations Breakdown",
       x = "Location",
       y = "Fraction of Population",
       fill = "Type of Vaccine") +
  coord_flip() + 
  theme_bw() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) 
```

Vaccination breakdown for all 50 states and DC.

```{r}
new_US_deaths_cases <- US_deaths_cases %>%
  filter(!(state %in% c("NYC", "PR", "GU",
                           "VI", "MP", "RMI",
                           "AS", "PW", "FSM"))) %>%
  group_by(submission_date) %>%
  summarise(total_new_cases = sum(new_case),
            total_new_deaths = sum(new_death),
            total_cases = sum(tot_cases),
            total_deaths = sum(tot_death)) 

new_US_deaths_cases$submission_date = as.Date(new_US_deaths_cases$submission_date,
                                              format="%m/%d/%Y")

new_US_deaths_cases <- new_US_deaths_cases %>%
  arrange(desc(submission_date))

new_US_deaths_cases
```

```{r}
new_US_deaths_cases %>% 
  slice_max(total_new_cases, n = 5) 

max_cases_row <- new_US_deaths_cases %>% 
  slice_max(total_new_cases, n = 1) 

max_cases_date <- max_cases_row$submission_date

max_cases_date
```

```{r}
new_US_deaths_cases %>%
  slice_max(total_new_deaths, n = 5)

max_deaths_row <- new_US_deaths_cases %>% 
  slice_max(total_new_deaths, n = 1) 

max_deaths_date <- max_deaths_row$submission_date

max_deaths_date
```

```{r}
as.integer(max_deaths_date) - as.integer(max_cases_date)
```

```{r}
x <- as.ts(c(11,7,8,12,14))
y <- as.ts(c(7,10,6,7,10))
x
```

```{r}
ccf_values1 = ccf(new_US_deaths_cases$total_new_cases, new_US_deaths_cases$total_new_deaths, lag.max = 100)
ccf_values1

# https://online.stat.psu.edu/stat510/lesson/8/8.2
```

```{r}
ccf_values2 = ccf(new_US_deaths_cases$total_cases, new_US_deaths_cases$total_deaths, lag.max = 100)
ccf_values2
```


```{r, fig.width=20, fig.height=12}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_new_cases)) +
  labs(title = "New Cases vs. Time",
       x = "Date",
       y = "Number of People") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r, fig.width=20, fig.height=12}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_new_deaths)) +
  labs(title = "New Deaths vs. Time",
       x = "Date",
       y = "Number of People") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r, fig.width=20, fig.height=12}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_cases / 10^6)) +
  labs(title = "Total Cases vs. Time",
       x = "Date",
       y = "Number of People (millions)") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r, fig.width=20, fig.height=12}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_deaths / 10^3)) +
  labs(title = "Total Deaths vs. Time",
       x = "Date",
       y = "Number of People (thousands)") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r}
new_US_vaccinations <- US_vaccinations %>%
  select(Date, Location, Series_Complete_Yes, 
         Series_Complete_Pop_Pct,
         Series_Complete_12PlusPop_Pct,
         Series_Complete_18PlusPop_Pct,
         Series_Complete_65PlusPop_Pct,
         Series_Complete_Janssen, Series_Complete_Moderna, 
         Series_Complete_Pfizer, Series_Complete_Unk_Manuf) %>%
  filter(Location %in% c("US"))

new_US_vaccinations$Date = as.Date(new_US_vaccinations$Date,
                                              format="%m/%d/%Y")

new_US_vaccinations
```

```{r, fig.width=20, fig.height=12}
new_US_vaccinations %>%
  filter(Series_Complete_Pop_Pct > 0) %>%
  ggplot() +
  geom_point(aes(x = Date, y = Series_Complete_Pop_Pct)) +
  labs(title = "Series Complete Vaccinations Pop. Pct vs. Time",
       x = "Date",
       y = "Percent of US Population") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r, fig.width=20, fig.height=12}
new_US_vaccinations %>%
  filter(Series_Complete_Yes > 0) %>%
  ggplot() +
  geom_point(aes(x = Date, y = Series_Complete_Yes / 10^6)) +
  labs(title = "Series Complete Vaccinations Total Pop. vs. Time",
       x = "Date",
       y = "Number of people (in millions)") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r, fig.width=20, fig.height=18}
Pop_Pct.labs <- c("% Fully Vaccinated 12+", 
                  "% Fully Vaccinated (Total Pop)",
                  "% Fully Vaccinated 18+", 
                  "% Fully Vaccinated 65+")
names(Pop_Pct.labs) <- c("Series_Complete_12PlusPop_Pct", 
                         "Series_Complete_Pop_Pct",
                         "Series_Complete_18PlusPop_Pct",
                         "Series_Complete_65PlusPop_Pct")

new_US_vaccinations %>%
  filter(Series_Complete_Yes > 0) %>%
  pivot_longer(cols = c(Series_Complete_Pop_Pct,
                        Series_Complete_12PlusPop_Pct,
                        Series_Complete_18PlusPop_Pct,
                        Series_Complete_65PlusPop_Pct),
               names_to = "Pop_Pct_Type",
               values_to = "Pop_Pct") %>%
  ggplot() +
  geom_point(aes(x = Date, y = Pop_Pct)) +
  facet_wrap(~Pop_Pct_Type,
             labeller = labeller(Pop_Pct_Type = Pop_Pct.labs)) + 
  labs(title = "Fully Vaccinated Pop. Pct vs. Time",
       x = "Date",
       y = "Percent of US Population") +
  theme_bw() +
  theme(strip.text = element_text(size = 20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

```{r}
merged_US_data <- inner_join(new_US_vaccinations, new_US_deaths_cases, by = c("Date" = "submission_date")) %>%
  filter(Series_Complete_Yes > 0)

merged_US_data
```

```{r}
ccf_values3 = ccf(merged_US_data$Series_Complete_Yes, merged_US_data$total_cases, lag.max = 150)
ccf_values3
```

```{r}
ccf_values4 = ccf(merged_US_data$Series_Complete_Yes, merged_US_data$total_deaths, lag.max = 250)
ccf_values4
```

```{r}
ccf_values4 = ccf(merged_US_data$Series_Complete_Pop_Pct, merged_US_data$total_new_deaths, lag.max = 150)
ccf_values4
```

```{r}
month(new_US_deaths_cases$submission_date)
```


```{r}
new_US_deaths_cases %>%
  mutate(season = case_when(
  month(submission_date) %in% c(9, 10, 11) ~ "Fall",
  month(submission_date) %in% c(12, 1, 2) ~ "Winter",
  month(submission_date) %in% c(3, 4, 5) ~ "Spring",
  month(submission_date) %in% c(6, 7, 8) ~ "Summer"
)) %>%
  group_by(season) %>%
  summarise(median_ncases = median(total_new_cases, na.rm = TRUE),
            median_ndeaths = median(total_new_deaths, na.rm = TRUE),
            mean_ncases = mean(total_new_cases, na.rm = TRUE),
            mean_ndeaths = mean(total_new_deaths, na.rm = TRUE),
            sd_ncases = sd(total_new_cases, na.rm = TRUE),
            sd_ndeaths = sd(total_new_deaths, na.rm = TRUE),
            min_ncases = min(total_new_cases, na.rm = TRUE),
            max_ncases = max(total_new_cases, na.rm = TRUE),
            iqr_ncases = IQR(total_new_cases, na.rm = TRUE))
```

```{r, fig.width=14, fig.height=10}
new_US_deaths_cases %>%
  mutate(season = case_when(
  month(submission_date) %in% c(9, 10, 11) ~ "Fall",
  month(submission_date) %in% c(12, 1, 2) ~ "Winter",
  month(submission_date) %in% c(3, 4, 5) ~ "Spring",
  month(submission_date) %in% c(6, 7, 8) ~ "Summer"
)) %>%
  ggplot(aes(x = total_new_cases / 10^3, 
             y = season, 
             fill = season)) +
  geom_density_ridges(alpha = 0.2) +
  labs(title = "Ridge plot of total new cases by season",
       x = "New Cases (Thousands)",
       y = "Season",
       fill = "Season") +
  theme_bw()
```

```{r, fig.width=14, fig.height=10}
new_US_deaths_cases %>%
  mutate(season = case_when(
  month(submission_date) %in% c(9, 10, 11) ~ "Fall",
  month(submission_date) %in% c(12, 1, 2) ~ "Winter",
  month(submission_date) %in% c(3, 4, 5) ~ "Spring",
  month(submission_date) %in% c(6, 7, 8) ~ "Summer"
)) %>%
  ggplot(aes(x = total_new_cases / 10^3, 
             y = season, 
             fill = season)) +
  geom_boxplot() +
  labs(title = "Ridge plot of total new cases by season",
       x = "New Cases (Thousands)",
       y = "Season",
       fill = "Season") +
  theme_bw()
```