---
title: "DCS210 Final Project"
subtitle: "Exploring COVID-19 Data"
author: "omarxinyijohn <br> John Mieszczanski & Omar Sarr & Xinyi Zhang"
institute: "Bates College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      color: #000000
---
```{r load-packages, include = FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(broom)
library(lubridate)
library(devtools)
library(dsbox)
library(ggridges)
library(openintro)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(reactable)
```

```{r setup, include = FALSE}
# For better figure resolution
#knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include = FALSE}
US_deaths_cases <- 
  read_csv("/cloud/project/data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")

US_vaccinations <- 
  read_csv("/cloud/project/data/COVID-19_Vaccinations_in_the_United_States_Jurisdiction.csv")
```

```{r, echo = FALSE}
style_xaringan(
  title_slide_background_color = "#000000",
  title_slide_text_color = "#FFFFFF"
)
```

### Introduction

Reported first in December, 2019, the COVID-19 pandemic constitutes the largest global public health crisis in a century, with daunting health and socioeconomic challenges. In the United States alone, more than 700,000 people have lost their lives due to the public health crisis. The progress in vaccination rates, as well as the spread of Delta variant, added more nuances to the overall situation.

### Overall Goal

In this project, we focus on the COVID-19 data in the U.S. We aim to explore:

- The fluctuations of deaths and cases across time since December, 2019
    
- The potential correlation between seasons and deaths and cases

- The potential correlation between vaccination rates and deaths and cases across time
    
- The differences in cases, deaths, and vaccinations across states

- The popularity of the main kinds of vaccines (J&J, Pfizer, Moderna) in different states

---

### COVID-19 Cases and Deaths - The Beginning
On January 22nd, 2020, the CDC began tracking daily COVID19 cases and deaths
across the United States and its territories. Eventually, on March 11th, 2020,
WHO declared COVID-19 a global pandemic. This day also coincided with the NBA
shutting down indefinitely. Two days later, Bates College announced they would 
be sending us home!

How do these two key dates compare the worst day(s) of the pandemic?
```{r, echo = FALSE}
new_US_deaths_cases <- US_deaths_cases %>%
  filter(!(state %in% c("NYC", "PR", "GU",
                           "VI", "MP", "RMI",
                           "AS", "PW", "FSM"))) %>%
  group_by(submission_date) %>%
  summarise(total_new_cases = sum(new_case),
            total_new_deaths = sum(new_death),
            total_cases = sum(tot_cases),
            total_deaths = sum(tot_death)) 

new_US_deaths_cases$submission_date = as.Date(new_US_deaths_cases$submission_date,
                                              format="%m/%d/%Y")

new_US_deaths_cases <- new_US_deaths_cases %>%
  arrange(desc(submission_date))

new_US_deaths_cases %>%
  filter(submission_date %in% c(as.Date("2020-01-22"), 
                                as.Date("2020-03-11"), 
                                as.Date("2020-03-13"), 
                                as.Date("2021-01-06"),
                                as.Date("2021-01-13"))) %>%
  reactable()
```

```{r, echo = FALSE, include = FALSE}
# How we obtained the date with the highest number of new cases
new_US_deaths_cases %>% 
  slice_max(total_new_cases, n = 5) 

max_cases_row <- new_US_deaths_cases %>% 
  slice_max(total_new_cases, n = 1) 

max_cases_date <- max_cases_row$submission_date

max_cases_date
```

```{r, echo = FALSE, include = FALSE}
# How we obtained the date with the highest number of new deaths
new_US_deaths_cases %>%
  slice_max(total_new_deaths, n = 5)

max_deaths_row <- new_US_deaths_cases %>% 
  slice_max(total_new_deaths, n = 1) 

max_deaths_date <- max_deaths_row$submission_date

max_deaths_date
```
---
<html>
<head>
<style>
p {margin-top: 0.5em ;
   margin-bottom: 1em ;}
h3 {margin-top: 0.5em ;
   margin-bottom: 0.5em ;}
</style>
</head>
<body>
<h3>COVID-19 Cases and Deaths - How did they evolve with time?</h3>
<p>Throughout this pandemic, we have experienced periods of low cases and deaths and periods of high cases and deaths. These ebbs and flows are typically called "waves". Our first goal was to visualize the "waves" of the pandemic - when were things the worst and when did things seem to be better?</p>
</body>
</html>

```{r, echo = FALSE, fig.width=14, fig.height=6}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_new_cases)) +
  labs(title = "Daily New Cases vs. Time",
       x = "Date",
       y = "Daily New Cases") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```
---
<html>
<head>
<style>
p {margin-top: 0.5em ;
   margin-bottom: 1em ;}
h3 {margin-top: 0.5em ;
   margin-bottom: 0.5em ;}
</style>
</head>
<body>
<h3>COVID-19 Cases and Deaths - How did they evolve with time?</h3>
</body>
</html>
```{r, echo = FALSE, fig.width=16, fig.height=8}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = submission_date, y = total_new_deaths)) +
  labs(title = "Daily New Deaths vs. Time",
       x = "Date",
       y = "Daily New Deaths") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```
---
### COVID-19 Cases and Deaths - What is the correlation?
Typically, when we are comparing two variables $X$ and $Y$, we calculate the 
correlation coefficient 
$$\rho = \frac{\sigma_{XY}}{\sigma_X\sigma_Y}$$ 
The correlation coefficient describes how one variable moves in relation to 
another. When dealing with time series data and both $X$ and $Y$ 
are functions of time, the time series $Y_t$ might be related to past lags of 
the time series $X_t$. In other words, if our time series is discrete 
(it must be), we want to determine the relationship between $X_{t + h}$ and
$Y_t$ for $h = 0, \pm 1, \pm 2, ...$. The **cross correlation coefficient** is 
thus given by
$$\rho = \frac{\sigma_{X_{t + h}Y_t}}{\sigma_{X_{t + h}}\sigma_{Y_t}}$$ 
Two things to note :

- When one or more $X_{t + h}$, with $h$ *negative*, are predictors of $Y_t$, 
it is sometimes said that **x leads y**

- When one or more $X_{t + h}$, with $h$ *positive*, are predictors of $Y_t$, 
it is sometimes said that **x lags y** $^*$

.footnote[
$^*$ https://online.stat.psu.edu/stat510/lesson/8/8.2
]
---
### COVID-19 Cases and Deaths - What is the correlation?
Before we use the R `ccf()` function, let's take a look at the relationship between
cases and deaths with no lag.

```{r, echo = FALSE, fig.width = 14, fig.height = 6}
new_US_deaths_cases %>%
  ggplot() +
  geom_point(aes(x = total_new_cases, y = total_new_deaths)) +
  labs(title = "Daily New Deaths vs. Daily New Cases",
       x = "Daily New Cases",
       y = "Daily New Deaths") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```
---
### COVID-19 Cases and Deaths - What is the correlation?
Before we use the R `ccf()` function, let's take a look at the relationship between
cases and deaths with no lag:

```{r, echo = FALSE}
deaths_cases_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(total_new_deaths ~ total_new_cases, data = new_US_deaths_cases) 
```

```{r, echo = FALSE, message = FALSE, fig.width = 14, fig.height = 6}
new_US_deaths_cases %>%
  ggplot(aes(x = total_new_cases, y = total_new_deaths)) +
  geom_point() +
  geom_line(aes(x = total_new_cases, y = augment(deaths_cases_fit$fit)$.fitted), 
            size = 0.75, 
            color = "blue") +
  labs(title = "Daily New Deaths vs. Daily New Cases",
       x = "Daily New Cases",
       y = "Daily New Deaths") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```
---
### COVID-19 Cases and Deaths - What is the correlation?
Results from the linear regression:
```{r, echo = FALSE}
deaths_cases_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(total_new_deaths ~ total_new_cases, data = new_US_deaths_cases) 

deaths_cases_fit

glance(deaths_cases_fit)
```

The linear model is given by `y = 0.01146x + 274.67437`. This tells us that for 
each additionally daily case, we expect the number of daily deaths to increase on 
average by `0.01146`. Additionally, we can see $R^2 = 0.673$, which indicates 
that only about 67.3% of the variation in the daily new deaths is dependent on 
daily new cases.
---
### COVID-19 Cases and Deaths - What is the correlation?
```{r, echo = FALSE, fig.width = 16, fig.height = 7.5}
xcf_plot <- function(df, x, y, title = "Cross Correlation", lag_var = 100){
  df_x <- eval(substitute(x), df)
  df_y <- eval(substitute(y), df)
  ccf.object <- ccf(df_x, df_y, plot = FALSE, lag.max = lag_var)
  output_table <- cbind(lag = ccf.object$lag, 
                        x.corr = ccf.object$acf) %>%
                          as_tibble() %>%
                          mutate(cat = ifelse(x.corr > 0, "green", "red"))
  output_table %>%
    ggplot(aes(x = lag, y = x.corr)) +
    geom_bar(stat = "identity", aes(fill = cat)) +
    scale_fill_manual(values = c("#339933", "#cc0000")) +
    ylab("Cross Correlation") +
    xlab("Lag (days)") +
    scale_y_continuous(limits = c(-1, 1)) +
    theme_bw() + 
    theme(legend.position = "none",
          axis.line = element_line(colour = "black"),
          axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
          axis.text.y = element_text(color = "grey20", 
                                     size = 15, 
                                     angle = 0, 
                                     hjust = .5, 
                                     vjust = .5, 
                                     face = "plain"),  
          axis.title.x = element_text(color = "grey20", 
                                      size = 30, 
                                      angle = 0, 
                                      hjust = .5, 
                                      vjust = 0, 
                                      face = "plain"),
          axis.title.y = element_text(color = "grey20", 
                                      size = 30, 
                                      angle = 90, 
                                      hjust = .5, 
                                      vjust = 1, 
                                      face = "plain"),
          plot.title = element_text(color = "grey20", 
                                      size = 30, 
                                      angle = 0, 
                                      hjust = .5, 
                                      vjust = .5, 
                                      face = "plain")) +
    ggtitle(title) -> p
}     

# run the function
plot1 <- xcf_plot(df = new_US_deaths_cases, 
         x = new_US_deaths_cases$total_new_cases, 
         y = new_US_deaths_cases$total_new_deaths, 
         title = "Cross Correlation : Daily New Cases and Daily New Deaths")

plot1
```
---
### COVID-19 Cases and Deaths - What is the correlation?
```{r, echo = FALSE, fig.width = 16, fig.height = 7.5}
plot2 <- xcf_plot(df = new_US_deaths_cases, 
         x = new_US_deaths_cases$total_new_cases, 
         y = new_US_deaths_cases$total_new_deaths, 
         title = "Cross Correlation : Daily New Cases and Daily New Deaths",
         lag_var = 10)

plot2
```
---
### COVID-19 Cases and Deaths - What is the correlation?
```{r, echo = FALSE, fig.show='hide'}
ccf_raw_data <- ccf(new_US_deaths_cases$total_new_cases, new_US_deaths_cases$total_new_deaths, plot = FALSE, lag.max = 14) 

ccf_raw_data <- data.frame(lag = ccf_raw_data$lag, CCF = ccf_raw_data$acf)
  
reactable(ccf_raw_data)
  
#data.frame(lag = correlation$ccf$lag, CCF = correlation$ccf$acf)
```
---
### COVID-19 Vaccinations Over Time
Now let's take a look at how the number people of vaccinated in the United States 
has evolved with time:
```{r, echo = FALSE}
new_US_vaccinations <- US_vaccinations %>%
  select(Date, Location, Series_Complete_Yes, 
         Series_Complete_Pop_Pct,
         Series_Complete_12PlusPop_Pct,
         Series_Complete_18PlusPop_Pct,
         Series_Complete_65PlusPop_Pct,
         Series_Complete_Janssen, Series_Complete_Moderna, 
         Series_Complete_Pfizer, Series_Complete_Unk_Manuf) %>%
  filter(Location %in% c("US"))

new_US_vaccinations$Date = as.Date(new_US_vaccinations$Date,
                                              format="%m/%d/%Y")
```

```{r, echo = FALSE, fig.width = 14, fig.height = 7}
new_US_vaccinations %>%
  filter(Series_Complete_Pop_Pct > 0) %>%
  ggplot() +
  geom_point(aes(x = Date, y = Series_Complete_Pop_Pct)) +
  labs(title = "Percent of US Population Fully Vaccinated vs. Time",
       x = "Date",
       y = "Percent of US Population") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```
---
### COVID-19 Vaccinations Over Time
Although the total population of the US is only about ~55% vaccinated as of early 
October, 2021, this does not paint a full picture. Let's break down this percentage
by age group:

```{r, echo = FALSE, fig.width = 14, fig.height = 7}
Pop_Pct.labs <- c("% Fully Vaccinated 12+", 
                  "% Fully Vaccinated (Total Pop)",
                  "% Fully Vaccinated 18+", 
                  "% Fully Vaccinated 65+")
names(Pop_Pct.labs) <- c("Series_Complete_12PlusPop_Pct", 
                         "Series_Complete_Pop_Pct",
                         "Series_Complete_18PlusPop_Pct",
                         "Series_Complete_65PlusPop_Pct")

new_US_vaccinations %>%
  filter(Series_Complete_Yes > 0) %>%
  pivot_longer(cols = c(Series_Complete_Pop_Pct,
                        Series_Complete_12PlusPop_Pct,
                        Series_Complete_18PlusPop_Pct,
                        Series_Complete_65PlusPop_Pct),
               names_to = "Pop_Pct_Type",
               values_to = "Pop_Pct") %>%
  ggplot() +
  geom_point(aes(x = Date, y = Pop_Pct)) +
  facet_wrap(~Pop_Pct_Type,
             labeller = labeller(Pop_Pct_Type = Pop_Pct.labs)) + 
  labs(title = "Percent of US Population Fully Vaccinated vs. Time",
       subtitle = "Broken Down by Age Group",
       x = "Date",
       y = "Percent of US Population") +
  theme_bw() +
  theme(strip.text = element_text(size = 20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),  
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain"),
        plot.subtitle = element_text(color = "grey20", 
                                    size = 20, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b %Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5))
```
---
### Exploring the Relationship Between Vaccinations and Cases/Deaths
Next, let's take a look at how vaccinations and deaths have evolved over time
in conjunction with one another:

```{r, echo = FALSE}
merged_US_data <- inner_join(new_US_vaccinations, new_US_deaths_cases, by = c("Date" = "submission_date")) %>%
  filter(Series_Complete_Yes > 0)
```

```{r, echo = FALSE, fig.width = 14, fig.height = 7}
coeff <- 1

merged_US_data %>%
  ggplot(aes(x = Date)) +
  geom_point(aes(y = total_deaths / 10^3), color = "red") + 
  geom_point(aes(y = Series_Complete_Yes / 10^6), color = "blue") +
  scale_color_manual(name = "Data", values = c("total_deaths" = "red", "Series_Complete_Yes" = "blue")) +
  scale_y_continuous(
    name = "Total # of COVID Deaths (Thousands)",
    sec.axis = sec_axis(~.*coeff, 
                        name = "Total # of People Fully Vaccinated (Millions)",
                        scales::pretty_breaks(n = 10)),
    breaks = scales::pretty_breaks(n = 10)
  ) +
  labs(title = "Total Fully Vaccinated and Total Deaths vs. Time",
       x = "Date",
       color = "Color") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.text.y = element_text(color = "grey20", 
                                   size = 15, 
                                   angle = 0, 
                                   hjust = .5, 
                                   vjust = .5, 
                                   face = "plain"),
        axis.title.x = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = 0, 
                                    face = "plain"),
        axis.title.y.left = element_text(color = "red", 
                                    size = 25, 
                                    angle = 90, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        axis.title.y.right = element_text(color = "blue", 
                                    size = 25, 
                                    angle = 270, 
                                    hjust = .5, 
                                    vjust = 1, 
                                    face = "plain"),
        plot.title = element_text(color = "grey20", 
                                    size = 30, 
                                    angle = 0, 
                                    hjust = .5, 
                                    vjust = .5, 
                                    face = "plain")) +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y")
```
---
### Exploring the Relationship Between Vaccinations and Cases/Deaths

---

## Number of Vaccinated People

--

## Proportion of Vaccinated People

--

## Differences across Age Groups

---

## COMBINE AND EXPLORE

--

# Have vaccines reduced cases and/or deaths and if so, to what extent?

---

## ADD ANOTHER DIMENSION

--

# Which vaccines are more popular in certain states (J&J, Pfizer, Moderna)?

---

# xaringan

- The presentation is created using the `xaringan` package

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
