---
title: "DCS210 Final Project"
subtitle: "Exploring COVID-19 Data"
author: "omarxinyijohn <br> John Mieszczanski & Omar Sarr & Xinyi Zhang"
institute: "Bates College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      color: #000000
---
```{r load-packages, include = FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(broom)
library(lubridate)
library(devtools)
library(dsbox)
library(ggridges)
library(openintro)
```

```{r setup, include = FALSE}
# For better figure resolution
#knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include = FALSE}
US_deaths_cases <- 
  read_csv("/cloud/project/data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")

US_vaccinations <- 
  read_csv("/cloud/project/data/COVID-19_Vaccinations_in_the_United_States_Jurisdiction.csv")
```

```{r, echo = FALSE}
style_xaringan(
  title_slide_background_color = "#000000",
  title_slide_text_color = "#FFFFFF"
)
```

### Introduction

Reported first in December, 2019, the COVID-19 pandemic constitutes the largest global public health crisis in a century, with daunting health and socioeconomic challenges. In the United States alone, more than 700,000 people have lost their lives due to the public health crisis. The progress in vaccination rates, as well as the spread of Delta variant, added more nuances to the overall situation.

### Overall Goal

In this project, we focus on the COVID-19 data in the U.S. We aim to explore:

- The fluctuations of deaths and cases across time since December, 2019
    
- The potential correlation between seasons and deaths and cases

- The potential correlation between vaccination rates and deaths and cases across time
    
- The differences in cases, deaths, and vaccinations across states

- The popularity of the main kinds of vaccines (J&J, Pfizer, Moderna) in different states

---

### COVID-19 Cases and Deaths - The Beginning
On January 22nd, 2020, the CDC began tracking daily COVID19 cases and deaths
across the United States and its territories. Eventually, on March 11th, 2020,
WHO declared COVID-19 a global pandemic. This day also coincided with the NBA
shutting down indefinitely. Two days later, Bates College announced they would 
be sending us home!

How do these two key dates compare the worst day(s) of the pandemic?
```{r, echo = FALSE}
new_US_deaths_cases <- US_deaths_cases %>%
  filter(!(state %in% c("NYC", "PR", "GU",
                           "VI", "MP", "RMI",
                           "AS", "PW", "FSM"))) %>%
  group_by(submission_date) %>%
  summarise(total_new_cases = sum(new_case),
            total_new_deaths = sum(new_death),
            total_cases = sum(tot_cases),
            total_deaths = sum(tot_death)) 

new_US_deaths_cases$submission_date = as.Date(new_US_deaths_cases$submission_date,
                                              format="%m/%d/%Y")

new_US_deaths_cases <- new_US_deaths_cases %>%
  arrange(desc(submission_date))

new_US_deaths_cases %>%
  filter(submission_date %in% c(as.Date("2020-01-22"), 
                                as.Date("2020-03-11"), 
                                as.Date("2020-03-13"), 
                                as.Date("2021-01-06"),
                                as.Date("2021-01-13"))) %>%
  kable(format = "html")
```

```{r, echo = FALSE, include = FALSE}
new_US_deaths_cases %>% 
  slice_max(total_new_cases, n = 5) 

max_cases_row <- new_US_deaths_cases %>% 
  slice_max(total_new_cases, n = 1) 

max_cases_date <- max_cases_row$submission_date

max_cases_date
```

```{r, echo = FALSE, include = FALSE}
new_US_deaths_cases %>%
  slice_max(total_new_deaths, n = 5)

max_deaths_row <- new_US_deaths_cases %>% 
  slice_max(total_new_deaths, n = 1) 

max_deaths_date <- max_deaths_row$submission_date

max_deaths_date
```
---
# How did the overall fluctuations in deaths & cases look like in the U.S.?

---

## Total Cases & New Cases

--

## Total Deaths & New Deaths

---

## BREAKDOWN BY SEASONS

--

# Are there any seasonal changes in cases & deaths?

---

## ZOOM IN BY STATES

--

# How do changes in cases & deaths differ across states?

---

## INTRODUCE ESSENTIAL FACTORS

--

# How did the overall vaccinations look like in the U.S.?

---

--

## Number of Vaccinated People

--

## Proportion of Vaccinated People

--

## Differences across Age Groups

---

## COMBINE AND EXPLORE

--

# Have vaccines reduced cases and/or deaths and if so, to what extent?

---

## ADD ANOTHER DIMENSION

--

# Which vaccines are more popular in certain states (J&J, Pfizer, Moderna)?

---

# xaringan

- The presentation is created using the `xaringan` package

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
